# Required libraries

```{r, echo=TRUE}
#install.packages("dplyr")
library(dplyr)
#install.packages("devtools")
library(devtools)
#install_github("easyGgplot2", "kassambara")
library(easyGgplot2)
library(ggplot2)
source("http://peterhaschke.com/Code/multiplot.R")
setwd("C:/Users/Pol/Documents/R")
```

# Code for reading in the dataset and/or processing the data

```{r, echo=TRUE}
data <- read.csv("C:/Users/Pol/Documents/R/activity.csv")
str(data)
dim(data)
```


# Histogram of the total number of steps taken each day

NA entries removed in the new dataset named **data_steps_yes** 

```{r, echo=TRUE}
data_steps_yes <- data[!is.na(data$steps),]
data_steps_na <- data[is.na(data$steps),]
nrow(data) == nrow(data_steps_yes) + nrow(data_steps_na)

data_steps_yes_perday <- group_by(data_steps_yes, date)
data_steps_yes_perday <- summarise(data_steps_yes_perday,sum_steps = sum(steps))
data_steps_yes_perday$date <- as.Date(data_steps_yes_perday$date)

ggplot(data_steps_yes_perday,aes(x=date,y=sum_steps)) + 
  geom_bar(colour="black", fill="#DD8888", width=.4,stat="identity") +
  guides(fill=FALSE) +
  xlab("Date") + ylab("Total Steps") +
  ggtitle("Histogram")
```  
  

# Mean and median number of steps taken each day

```{r, echo=TRUE}
data_steps_yes_perday_M <- group_by(data_steps_yes, date)
data_steps_yes_perday_M <- summarise(data_steps_yes_perday_M,M_steps = mean(steps))
data_steps_yes_perday_M$date <- as.Date(data_steps_yes_perday_M$date)

data_steps_yes_perday_MD <- group_by(data_steps_yes, date)
data_steps_yes_perday_MD <- summarise(data_steps_yes_perday_MD,MD_steps = median(steps))
data_steps_yes_perday_MD$date <- as.Date(data_steps_yes_perday_MD$date)

summary(data_steps_yes_perday_M)
head(data_steps_yes_perday_M)

summary(data_steps_yes_perday_MD)
head(data_steps_yes_perday_MD)

```  


# Time series plot of the average number of steps taken
```{r, echo=TRUE}
data_steps_yes_interval_M <- group_by(data_steps_yes, interval)
data_steps_yes_interval_M <- summarise(data_steps_yes_interval_M,M_int = mean(steps))

ggplot(data_steps_yes_interval_M,aes(x=interval,y=M_int)) + 
  geom_line(colour="gray55", linetype="solid", size=1) + 
    xlab("Interval") + ylab("Average Steps") +
  ggtitle("Time Series") 
```  


# The 5-minute interval that, on average, contains the maximum number of steps

```{r, echo=TRUE}
data_steps_yes_interval_M_max <- data_steps_yes_interval_M[data_steps_yes_interval_M$M_int == max(data_steps_yes_interval_M$M_int),]
data_steps_yes_interval_M_max
```
Interval with max number of steps is **`r data_steps_yes_interval_M_max[1]`**



# Code to describe and show a strategy for imputing missing data

Use **data_steps_yes_interval_M** dataset for intervals with missing data and later append results to the  original dataset **data_steps_yes** where NA has been removed 

```{r, echo=TRUE}
data_steps_na_filled <- merge(data_steps_na,data_steps_yes_interval_M,by = "interval")
data_steps_na_filled <- data_steps_na_filled[-c(2)]
data_steps_na_filled <- data_steps_na_filled[c(3,2,1)]
colnames(data_steps_na_filled)[1] <- "steps"

data_steps_all_filled <- rbind(data_steps_yes,data_steps_na_filled)
nrow(data_steps_all_filled) == nrow(data)
```




# Histogram of the total number of steps taken each day after missing values are imputed
```{r, echo=TRUE}
data_steps_all_filled_s <- group_by(data_steps_all_filled, date)
data_steps_all_filled_s <- summarise(data_steps_all_filled_s,sum_steps = sum(steps))
data_steps_all_filled_s$date <- as.Date(data_steps_all_filled_s$date)

ggplot(data_steps_all_filled_s,aes(x=date,y=sum_steps)) + 
  geom_bar(colour="black", fill="#DD8888", width=.4,stat="identity") +
  guides(fill=FALSE) +
  xlab("Date") + ylab("Total Steps") +
  ggtitle("Histogram")
```



# Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
data_steps_all_filled_M <- group_by(data_steps_all_filled, date)
data_steps_all_filled_M <- summarise(data_steps_all_filled_M,M_steps = mean(steps))
data_steps_all_filled_M$date <- as.Date(data_steps_all_filled_M$date)

data_steps_all_filled_MD <- group_by(data_steps_all_filled, date)
data_steps_all_filled_MD <- summarise(data_steps_all_filled_MD,MD_steps = median(steps))
data_steps_all_filled_MD$date <- as.Date(data_steps_all_filled_MD$date)

Mean_original_dataset <- summary(data_steps_yes_perday_M)
Median_original_dataset <- summary(data_steps_yes_perday_MD)

Mean_filled_dataset <- summary(data_steps_all_filled_M)
Median_filled_dataset <- summary(data_steps_all_filled_MD)


```

Interval with max number of steps is **`r data_steps_yes_interval_M_max[1]`**

**Original dataset**

**Mean:** `r Mean_original_dataset`
**Median:** `r Median_original_dataset`

**Filled-in dataset**

**Mean:** `r Mean_filled_dataset`
**Median:** `r Median_filled_dataset`

No impact to mean values

Median values are **higher** in filled-in dataset


# Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


```{r, echo=TRUE}
data_steps_all_filled_daytype <- transform(data_steps_all_filled,flag_weekend = (weekdays(as.Date(date)) %in% c('Saturday','Sunday')))
data_steps_all_filled_Week <- data_steps_all_filled_daytype[data_steps_all_filled_daytype$flag_weekend==TRUE,]
data_steps_all_filled_Weekend <- data_steps_all_filled_daytype[data_steps_all_filled_daytype$flag_weekend==FALSE,]


data_steps_all_filled_Weekend_M <- group_by(data_steps_all_filled_Weekend, interval)
data_steps_all_filled_Weekend_M <- summarise(data_steps_all_filled_Weekend_M,M_int = mean(steps))

g1 <- ggplot(data_steps_all_filled_Weekend_M,aes(x=interval,y=M_int)) + 
  geom_line(colour="gray55", linetype=1, size=1) + 
  #geom_point(colour="green", size=1, shape=16) +
  xlab("Interval") + ylab("Average Steps") +
  ggtitle("Weekend") +
  scale_y_continuous(limits = c(0, 250))



data_steps_all_filled_Week_M <- group_by(data_steps_all_filled_Week, interval)
data_steps_all_filled_Week_M <- summarise(data_steps_all_filled_Week_M,M_int = mean(steps))

g2<- ggplot(data_steps_all_filled_Week_M,aes(x=interval,y=M_int)) + 
  geom_line(colour="gray55", linetype=1, size=1) + 
  #geom_point(colour="green", size=1, shape=16) +
  xlab("Interval") + ylab("Average Steps") +
  ggtitle("Weekdays") +
  scale_y_continuous(limits = c(0, 250))

multiplot(g1,g2, cols=2)
```

  
